---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    keep_md: yes
---



**Download data and put it into a data frame** 


```{r}
fileurl <- "http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(fileurl, "repdata-data-activity.zip", mode="wb")
unzip("repdata-data-activity.zip")
activity <- read.csv("activity.csv")

```
***What is mean total number of steps taken per day?***

```{r}
## Total the steps
Total_steps <- aggregate(steps ~ date, data = activity, sum)

## Set labels to variables
MyMainLabel <- "Total steps by day";  MyXLabel <- "Steps per day"; MyYLabel <- "Number of times"

## Show on Histogram
hist(Total_steps$steps, main = MyMainLabel , xlab = MyXLabel, ylab = MyYLabel,col="green")
```


**Mean of Total Steps** 

```{r}
mean(Total_steps$steps)
```

**Median of Total Steps** 

```{r}
median(Total_steps$steps)
```

***What is the average daily activity pattern?***

```{r}
## Get mean of steps
stepsMean <- aggregate(data=activity,steps~interval,FUN=mean,  na.action=na.omit)

## Set labels to variables
TSmainLabel <- "Average Daily Activity Pattern";  TSXLabel <- "5-min Interval"; TSYLabel <- "Average Num of Steps"

## Plot in time series
plot(stepsMean$interval, stepsMean$steps,type = "l",
     xlab = TSXLabel, ylab = TSYLabel,main = TSmainLabel)
```


```{r}
# Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
stepsMean$interval[which.max(stepsMean$steps)]
```

***Imputing missing values***

```{r}
# Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
naValues <- sum(is.na(activity$steps)); naValues
```

```{r}
# Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. 

# Strategy put original data into new data frame and loop in mean value into NA's

activityNoNA <- activity
for (i in which(sapply(activityNoNA, is.numeric))) {
  activityNoNA[is.na(activityNoNA[, i]), i] <- mean(activityNoNA[, i],  na.rm = TRUE)
}
# Check to see if any NA's are left
sum(is.na(activityNoNA$steps))
# "activityNoNA" is a new data set with missing data filled in
```

```{r}
## Make a histogram of the total number of steps taken each day for the new data set
Total_stepsNoNA <- aggregate(steps ~ date, data = activityNoNA, sum)

hist(Total_steps$steps, main = MyMainLabel , xlab = MyXLabel, ylab = MyYLabel,col="orange")
```

**Mean and Median of Total Steps without missing values** 

```{r}
mean(Total_stepsNoNA$steps)
median(Total_stepsNoNA$steps)
```
Do these values differ from the estimates from the first part of the assignment? 

*  Yes, the values for mean and median are now equal.

What is the impact of imputing missing data on the estimates of the total daily number of steps?

*  The imputed mean values for missing values cause the median value to shift to the mean value.

**Are there differences in activity patterns between weekdays and weekends?** 
```{r}
# Create a new factor variable in the new dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

# First change date values into date format
activityNoNA$date<-as.Date(activityNoNA$date, format = '%Y-%m-%d')

# Then create part of week date belongs to as weekend or weekday and make factor
activityNoNA$weekPart <- ifelse(weekdays(activityNoNA$date) %in%  c("Saturday", "Sunday"),'weekend','weekday')
activityNoNA$weekPart <- as.factor(activityNoNA$weekPart)
```
```{r}
# Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 

plotData <- aggregate(steps ~ interval + weekPart, data=activityNoNA, mean)
library("lattice")
xyplot(type="l",data=plotData,steps ~ interval | weekPart,
  xlab=TSXLabel,ylab=TSYLabel,layout=c(1,2))
```